---
title: "In-class Exercise 5"
description: |
  Time-series analysis with R
author:
  - name: Antonius Handy 
    url: https://www.linkedin.com/in/antoniushandy
    affiliation: Singapore Management University, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

# 1. Getting Started

## Setting up R packages

To start with, **lubridate, tidyverse, data.table, ggthemes and ViSiElse** packages will be launched using `library()`.

```{r}
packages = c('lubridate', 'tidyverse','data.table','ggthemes','ViSiElse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}
```

# 2. Data Preparation

## Importing Data

```{r}
attacks <- read_csv("data/eventlog.csv")
```

## Data Wrangling

Two new fields namely *wkday* and *hour* will be derived using these two steps below.

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')
attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

# 3. Calendar Heatmap

## Single Heatmap

```{r}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()
ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

## Multiple Heatmaps

We are going to identify the top 4 countries with the highest number of attacks.

# 4. Working with ViSiElse

```{r}
data("typDay")
```

By default, ViSiElse shows the summary statistics (25% percentile, 50% percentile and 75% percentile). It purposely locate the circles far away from the blue bars to avoid overlapping. The distance between the first and third circle shows how large the IQR (Inter Quartile Range) is.

```{r}
visielse(typDay)
```

To omit this summary statistics, we can execute the below code chunk.

```{r}
visielse(typDay, informer = NULL)
```

The code chunk below will replace median by mean.

```{r}
visielse(typDay, informer = "mean")
```

Next, we will differentiate actions into two categories namely punctual and long. A punctual action is an action with no duration or not lasting long enough. A long action is an action having duration.

```{r}
p1 <- visielse(typDay, informer = NULL)
b1 <- ConvertFromViSibook(p1@book)
```

```{r}
b1 <- b1[order(as.numeric(b1$showorder)), ] # order the data.frame 
b1$label <- c("Sleep", "Stop sleeping", "Wake up", "Take a shower", "Eat breakfast", 
              "Start working", "Start eating lunch", "End of lunch", 
              "Stop working", "Pick up the kids", "Start cooking", 
              "End of dinner", "Go to sleep", "First coffee")
b1[15,] <- c("sleep", "Sleeping", "l", 1, "start_sleep", "stop_sleep")
b1[16,] <- c("work", "Working", "l", 5, "start_work", "stop_work")
b1[17,] <- c("lunch", "Lunch break", "l", 6, "start_lunch", "stop_lunch")
b1[18,] <- c("cook", "Cook and eat dinner", "l", 8, "start_cook", "stop_cook")
b1$showorder <- c(NA, NA, 2, 3, 4, 5, NA, NA, 7, 9, NA, NA, 11, 12, 1, 6, 8, 10) 
b1 <- b1[order(as.numeric(b1$showorder)), ]
```

The shape of long actions is different from punctual actions because 100 people have different start time and stop time.

```{r}
v2 <- visielse(typDay, book = b1, informer = NULL, doplot = F, pixel = 30)
plot(v2, vp0w = 0.7, unit.tps = "min", scal.unit.tps = 30, main = "Typical day")
```
