---
title: "Take-home Exercise 6"
description: |
  Community Network Visualisation
author:
  - name: Antonius Handy 
    url: https://www.linkedin.com/in/antoniushandy
    affiliation: Singapore Management University, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# 1. Overview

In this take-home exercise 6, we are going to reveal the patterns of **community interactions** of the city of Engagement, Ohio USA by using appropriate social network analysis approach.

The data is processed by using appropriate tidyverse family of packages, whereas the statistical graphics are prepared using ggraph/visNetwork and its extensions.

# 2. Getting Started

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The packages required for this exercise are **tidyverse**, **tidygraph**, **ggraph**, **visNetwork**, **lubridate** and **clock**.

```{r}
packages = c('tidyverse','tidygraph','ggraph','visNetwork','lubridate','clock')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 3. Data

### 3.1 Data Source
The original datasets were obtained from [*VAST Challenge 2022*](https://vast-challenge.github.io/2022/index.html) in csv format. They show the social interaction among 1011 residents of Engagement, OH that have agreed to participate in this study.

### 3.2 Importing Data


### 3.3 Data Wrangling


# 4. Social Network Visualisation


# 5. References





