---
title: "Take-home Exercise 4"
description: |
  Daily Routines Visualisation
author:
  - name: Antonius Handy 
    url: https://www.linkedin.com/in/antoniushandy
    affiliation: Singapore Management University, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# 1. Overview

In this take-home exercise 4, we are going to analyse and desribe the **daily routines** of two participants of the city of Engagement, Ohio USA by using appropriate visual analytics methods.

The data is processed by using appropriate tidyverse family of packages, whereas the statistical graphics are prepared using ViSiElse and its extensions.

# 2. Getting Started

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The packages required for this exercise are [**tidyverse**](https://www.tidyverse.org/) and [**ViSiElse**](https://cran.r-project.org/web/packages/ViSiElse/index.html).

```{r}
packages = c('tidyverse','ViSiElse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 3. Data

### 3.1 Data Source
The original dataset was obtained from [*VAST Challenge 2022*](https://vast-challenge.github.io/2022/index.html) in csv format. It shows the daily routines of 1011 residents of Engagement, OH that have agreed to participate in this study from March 2022 until May 2023.

### 3.2 Importing Data

### 3.3 Data Wrangling


# 4. Daily Patterns

# 5. References