---
title: "Take-home Exercise 2"
description: |
  Creating data visualisation beyond default
author:
  - name: Antonius Handy 
    url: https://www.linkedin.com/in/antoniushandy
    affiliation: Singapore Management University, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# 1. Overview

In this take-home exercise, we are going to reveal the demographic of the city of Engagement, Ohio USA by using appropriate static statistical graphics methods.

The data is processed by using appropriate tidyverse family of packages, whereas the statistical graphics are prepared using ggplot2 and its extensions.

# 2. Getting Started

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The packages required for this exercise are [**tidyverse**](https://www.tidyverse.org/), [**ggdist**](https://mjskay.github.io/ggdist/index.html) and [**hrbrthemes**](https://www.rdocumentation.org/packages/hrbrthemes/versions/0.1.0).

```{r}
packages = c('tidyverse','ggdist','hrbrthemes')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 3. Data

### DATA SOURCE
The original dataset was obtained from [*VAST Challenge 2022*](https://vast-challenge.github.io/2022/index.html) in csv format. It consists of basic information about the residents of Engagement, OH that have agreed to participate in
this study.

### IMPORTING DATA
The code chunk below imports dataset into R by using `read_csv()` of **readr** and saves it as a tibble data frame called *demographics*. It consists of 1011 records as shown below.

```{r}
demographics <- read_csv("data/Participants.csv")
glimpse(demographics)
```




