---
title: "Take-home Exercise 3"
description: |
  Creating data visualisation beyond default
author:
  - name: Antonius Handy 
    url: https://www.linkedin.com/in/antoniushandy
    affiliation: Singapore Management University, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# 1. Overview

In this take-home exercise 3, we are going to reveal the **economic** of the city of Engagement, Ohio USA by using appropriate static and interactive statistical graphics methods.

The data is processed by using appropriate tidyverse family of packages, whereas the statistical graphics are prepared using ggplot2 and its extensions.

# 2. Getting Started

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The packages required for this exercise are [**tidyverse**](https://www.tidyverse.org/).

```{r}
packages = c('tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 3. Data

### 3.1 Data Source
The original datasets were obtained from [*VAST Challenge 2022*](https://vast-challenge.github.io/2022/index.html) in csv format. It consists of basic information about the residents of Engagement, OH that have agreed to participate in
this study.

### 3.2 Importing Data


### 3.3 Data Wrangling



# 4. Economic Visualisation

# 5. References
